Unifying quasars with clumpy wind models

### 2015-09-01 paper.tex
Knox's first set of changes, trying to commit them again

%% declare any packages used
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{color}
\usepackage{pdfpages}
\usepackage{appendix}
\usepackage{subfigure}
\usepackage{url}
%\usepackage{pngfig}
%\usepackage[dvips]{color}
%\usepackage{aabib}

@@

%% Journal definitions
\def\araa{ARAA}
\def\nat{Nature}
\def\apjl{ApJ Letters}
\def\aapr{AAPR}
\def\ssr{SSR}
\def\apj{ApJ}
\def\apjs{ApJs}
\def\pasp{PASP}
\def\aap{A\&A}
\def\mnras{MNRAS}
\def\aj{AJ}
\def\rmxaa{RMXAA}
\def\aaps{A\&As}

@@

%\author[Matthews et al.]
\author{
\parbox[t]{\textwidth}{
J.~H.~Matthews$^1$\thanks{jm8g08@soton.ac.uk}, C.~Knigge$^1$,
N.~Higginbottom$^1$, K.~S.~Long$^2$, S.~A.~Sim$^3$ and S.~W.~Mangham$^1$
}
\medskip  
\\$^1$School of Physics and Astronomy, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom
\\$^2$Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD, 21218
\\$^3$School of Mathematics and Physics, Queens University Belfast, University Road, Belfast, BT7 1NN, Northern Ireland, UK
}

@@

Various unification schemes for quasars and luminous active galactic nuclei (AGN) 
propose that much of their complex phenomenology can be explained by a simple 
geometrical picture involving an accretion disc and associated outflow.
Here, we test this paradigm by utilising our state-of-the-art radiative transfer 
and photoionization code to produce synthetic spectra from simple biconical disc wind models. 
In particular, we expand on our previous work by presenting a model that
exhibits many of the spectral features expected from a quasar unification model.

@@

allows for a more realistic X-ray luminosity in the model,
while maintaining the ionization state necessary for strong BAL features. 
We examine the X-ray properties of this new model

@@

We find that the dense, X-ray heated wind 

@@

The highest inclination models possess prominent \mg\ and \al\ BALs, the absorption 

@@

seen in BAL and non-BAL quasar composites.

@@

@@

The spectra of 
quasars and luminous active galactic nuclei (AGN) 
typically exhibit a series of strong emission lines
with an underlying blue continuum - the so-called {\sl `big blue bump'} (BBB). 
The BBB is often attributed to emssion from a 
geometrically thin, optically thick accretion disc surrounding the central black hole, similar
to that described by \cite{shakurasunyaev1973}.
In addition to the inflowing accreting material, i
outflows are ubiquitous in AGN
and quasars \citep{kellerman1989,ganguly2008}. These outflows can take the form of 
highly collimated radio jets \citep[e.g.][]{hazard1963,potash1980,perley1984,marscher2006}, 
or mass-loaded `winds' emanating from the accretion disc 
\citep{weymann1991,turnermiller2009}. 
Outflows in AGN offer a 
potential feedback mechanism through which the central source can 
affect its environment \citep{king2003,king2005,fabian2012}
-- feedback that is required in models of galaxy evolution \citep{springel2005}
and may explain the `$M-\sigma$' relation \citep{silkrees1998,haring2004}.

@@

Despite the clear importance of ADWs in understanding AGN SEDs and accretion physics, 
much of the underlying outflow physics remains highly uncertain. 
Several possible driving mechanisms for ADWs have been proposed, including
thermal pressure \citep{weymann1982, begelman1991}, magnetocentrifugal forces 
\citep{blandfordpayne,pelletier_pudritz} and 
radiation pressure on spectral lines \citep[`line-driving';][]{lucysolomon1970,shlosman1985,MCGV95}.
Of these, line-driving is possibly the most attractive, as
strong absorption lines are already seen in BALQSOs and the X-ray spectra of AGN 
\citep{reeves2003,poundsreeves2009,tombesi2010a}.
The presence of line-locked features \citep{bowler2014} 
and the `ghost of \la' (Arav et al. 1996; Arav 1996; North 2006; but see 
also Cottis et al. 2010) \nocite{arav1995, arav1996, north2006,cottis2010}
in the spectra of some BALQSOs also gives clearer evidence that line-driving is
at least partially contributing to the acceleration of the wind.

@@

There has been some success using simple
kinematic prescriptions for biconical disc winds to model 
AGN and quasar outflows \citep[][hereafter H13]{simlong2008,sim2010,higginbottom2013}. 
H13 successfully produced a benchmark model for BALQSOs. However, the model
had two key drawbacks. First, an unrealistically low X-ray luminosity
was required in order to prevent over-ionization of the outflow.
Second, the model failed to produce the strong emission lines 
required at low inclinations in a unified model.
In this paper, we attempt to address these issues, and test the disc wind 
unification model using Monte Carlo radiative transfer (MCRT) and photoionization 
calculations. The paper is organised as follows.
In section 2, we describe some of the important photoionization 
and MCRT aspects of the code. In section 3, we outline the model, including 
a description of our clumping implementation. In section 4, we present the results 
from a clumped model. In section 5 we discuss our results, 
focussing particular on the anisotropy of 
disc emission and GR effects, and finally, in section 6, we summarise our findings.

@@

We use the MCRT code \py\ to carry out our radiative transfer and photoionization
calculations in non-local thermodynamic equilibrium (non-LTE). 
The code has been used to model accreting white dwarfs 
(Long \& Knigge 2002, hereafter LK02; Noebauer et al. 2010; 
Matthews et al. 2015, hereafter M15), young-stellar objects \citep{simmacro2005}
and quasars (H13). \nocite{noebauer, M15, LK02}
These studies contain extensive detail on the code, 
so we only briefly describe the key elements of the global 
ionization calculation and other important aspects.

@@

To treat line transfer, we adopt the same hybrid scheme 
described by M15, 
in which the energy flows
through the system are described in terms of indivisible
energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section 2.3).
These energy packets interact with either two-level `simple ions'
or full `macro-atoms'. The macro-atom implementation 
is described in full by \cite{lucy2002, lucy2003}.
Our scheme allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements which are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions (see M15). In this study,
only H \& He are treated as macro-atoms, because 
we expect recombination to be important
in determining their level populations and resultant line emission.

@@

Macro-atoms have their ion and level populations derived from
MC rate estimators as described by M15. Previously (LK02, H13, M15),
we adopted a modified Saha approach to calculate the ionization fractions
of simple-ions. We have now improved our code to explicitly solve the 
rate equations between ions in non-LTE. 
This dispenses with a number of small assumptions made in the modified Saha approach, 
is more numerically stable, 
and, in principle, allows the direct addition of extra physical 
processes such as Auger ionization.

%To avoid approximating the complex AGN SED as e.g. a dilute blackbody \citep[][M15]{ML93},
We model the SED in a cell using the technique described by H13. In this scheme,
the mean intensity in a series of $n$ bands is modeled as a normalised power law in 
frequency $\nu$
\begin{equation}
J_{\nu,i}=K_{pl}\nu^{\alpha_{pl}},
\end{equation}
for a band $i$, or an exponential 
\begin{equation}
J_{\nu,i}=K_{exp} e^{(-h\nu/k T_{exp})}.
\end{equation}
Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
deduced from the band-limited radiation field estimators.
The ionization rate out of ion $j$ can then be written as 
\begin{equation}
R_{j,j+1}(J)= 
\displaystyle{
n_i \left(C_{j} n_e + 
\sum_{band~i=0}^{n}~
{\int_{\nu_i}^{~\nu_i+1}{J_{\nu,i}\sigma_j(\nu)\nu^{-1}d\nu}}
\right),}
\end{equation}
where $\sigma_j$ is the photoionization cross-section and $C_{j}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $j$ is given by 
\begin{equation}
R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{DR} + \alpha^j_{CR}) n_{j+1} n_e,
\end{equation}
where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
The subscripts denote radiative, dielectronic and 
collisional (three-body) recombination, respectively.
We neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann equation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$.

@@

@@

@@

\subsection{Synthetic Spectra}

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/windnew3.png}
\caption
{
A cartoon describing the broad classes of sightline 
in our model, illustrates how geometric effects lead to 
the different emergent spectra. The colour gradient is approximate,
but indicates the stratified ionization structure, 
from highly ionized (yellow) to low ionization (purple) material.
%[FIGURE NEEDS IMPROVING]
}
\label{fig:sightline}
\end{figure}

@@

Figure~\ref{fig:uvspec} shows the synthetic spectrum in the UV from our model. 
We also show a comparison to composite quasar and BALQSO spectra.
We show a cartoon illustrating how geometric effects determine
the output spectra in figure~\ref{fig:sightline}.

@@

We find that the model can produce significant line emission
at low inclinations, particular in \civ, 
and the improved treatment of recombination results in a strong \la\
line. In the context of unification, this is a promising result, 
and shows that a biconical wind can produce significant 
emission at `quasar-like' angles. To assess the ability of the model to match real 
quasar spectra, we also show {\sl Sloan Digital Sky Survey} (SDSS) quasar composites from
\cite{reichard2003}, normalised to the flux at 2000\AA\ in each panel.
We do not produce the semi-forbidden 
intercombination lines seen in quasar spectra because we
currently do not have a treatment for semi-forbidden lines.
This is especially noticable with the strong C~\textsc{iii}]~1909\AA\, 
line in the quasar composite spectra.
The critical density of the C~\textsc{iii}]~1909\AA\, line 
is $n_e\sim10^{9.5}~\rm{cm^{-3}}$, which is higher than much of the 
outer portion of our wind. We therefore expect intercombination lines
to become important coolants in the outer portion of the wind.

@@

@@

When comparing BALQSO and quasar composites, it is apparent
that they possess remarkably similar line strengths and widths 
\citep[e.g.][]{reichard2003}.
This presents a challenge to our model, as well as the geometric 
unification picture in general.
Limb darkening and foreshortening causes the disc continuum emission to be 
strongly anisotropic, yet the line emission in our models is much more isotropic. 
This has the effect of enhancing the line-to-continuum ratios at high inclinations. 
To construct a scenario where emission 
line equivalent widths are comparable at all inclinations requires 
significant fine-tuning when considering a classical thin disc. 
An alternative, simpler solution is that the emergent continuum is roughly isotropic.

General relativistic effects -- specifically, light bending
and relativistic beaming -- can cause 
the accretion disc SED to become more isotropic \citep[e.g.][]{zhang1997,munozdarias2013}.
To generate GR disc spectra, we use the code \agn\ \citep{hubeny2000,davishubeny2006,davis2007}. 
The output flux at three different wavelengths
as a function of inclination for an \agn\ model with the same disc and BH parameters
as our clumpy wind model is shown in figure~\ref{fig:f2000}.
The effects of GR on an AGN disc are much less extreme 
in the UV portion of the spectrum than the calculations
by \cite{zhang1997} for the X-rays in X-ray binaries.
As a result, the disc emission is still strongly anisotropic.
GR alone therefore cannot explain the line ratio trends in quasars by making the disc
emit more isotropically.
This supports the findings of \cite{risaliti2011}, who find that EW distributions 
in quasars are consistent with anisotropic emission from optically thick, 
disc-like structures for {\em both} the continuum source and BLR. 
We will explore these ideas further in a future study.

@@

@@

>>>>

%% declare any packages used
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{color}
\usepackage{pdfpages}
\usepackage{appendix}
\usepackage{subfigure}
\usepackage{url}
%\usepackage{pngfig}
%\usepackage[dvips]{color}mak
%\usepackage{aabib}

@@

%% Journal definitions
\def\araa{ARAA}
\def\nat{Nature}
\def\apjl{ApJ Letters}
\def\aapr{AAPR}
\def\ssr{SSR}
\def\apj{ApJ}
\def\apjs{ApJs}
\def\pasp{PASP}
\def\aap{A\&A}
\def\mnras{MNRAS}
\def\aj{AJ}
\def\rmxaa{RMXAA}
\def\aaps{A\&As}
\def\LA{Lyman\thinspace$\alpha$}

\newcommand{\EXPN}[2]{\mbox{$#1\times 10^{#2}$}}
\newcommand{\EXPU}[3]{\mbox{\rm $#1 \times 10^{#2} \rm\:#3$}}  % exponent with units
\newcommand{\POW}[2]{\mbox{$\rm10^{#1}\rm\:#2$}}
\def\LUM{\:{\rm ergs\:s^{-1}}}
\def\FLUX{\:{\rm ergs\:cm^{-2}\:s^{-1}}}
\def\OIGS{\:{\rm ergs\:cm^{-2}\:s^{-1}\:\AA^{-1}}}

@@

%\author[Matthews et al.]
\author{
\parbox[t]{\textwidth}{
James~H.~Matthews$^1$\thanks{jm8g08@soton.ac.uk}, Christian~Knigge$^1$,
Nick~Higginbottom$^1$, Knox~S.~Long$^2$, Stuart~A.~Sim$^3$ and Sam~W.~Mangham$^1$
}
\medskip  
\\$^1$School of Physics and Astronomy, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom
\\$^2$Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD, 21218
\\$^3$School of Mathematics and Physics, Queens University Belfast, University Road, Belfast, BT7 1NN, Northern Ireland, UK
}

@@

Various unification schemes have been proposed to interpret the complex phenomenology of quasars and luminous active galactic nuclei (AGN) in terms of  a simple axisymmetric picture involving a central black hole, an accretion disc and an associated outflow.
Here, we continue our tests of  this paradigm by comparing ``average'' AGN spectra to synthetic spectra of simple biconical disc wind models, using  with our Monte Carlo radiative transfer and photoionization code {\sc Python}.
In particular, we investigate whether clumping of the outflow allows us to produce synthetic spectra in the rest-frame UV that have the characteristics of AGN, assuming total luminosities as high as 10$^{45}$ ergs s$^{-1}$. 

@@

allows for a more realistic AGN luminosities,
while maintaining the ionization state necessary for strong BAL features in the rest-frame UV. 
We examine the X-ray properties of these simple clumped models

@@

The dense, X-ray heated wind 

@@

At the highest inclinations, the synthetic spectra possess prominent \mg\ and \al\ BALs, the absorption 

@@

seen in BAL and non-BAL quasar composites. {\bf This sentence needs to be more precise, especially in light ot the statement earlier in the abstract about ``good agreement''}

@@

{\bf There is confusion in the abstract, at least, in what is the model and what is the synthetic spectra, as well as whether there is one model or many models.  We need to try to be clear.  The model should be a clumped biconical outflow.  There are then particular instantiations of the model. And then there are synthetic spectra}

@@

The spectra of 
quasars and luminous active galactic nuclei (AGN) 
typically exhibit a series of strong emission lines
with an underlying blue continuum - the so-called {\sl `big blue bump'} (BBB). 
The BBB is often attributed to emssion from a 
geometrically thin, optically thick accretion disc surrounding the central black hole, similar
to that described by \cite{shakurasunyaev1973}.  The lines are usually attributed to scattering and absorption of material above the disk.  {\bf Note: the first half of this paragraph and the second do not follow from one another.  The first half is about spectra the second about outflows and the connection to spectra is not made until the next paragrph.  This needs fixing.}
In addition to the inflowing accreting material, 
outflows are ubiquitous in AGN
and quasars \citep{kellerman1989,ganguly2008}. These outflows can take the form of 
highly collimated radio jets \citep[e.g.][]{hazard1963,potash1980,perley1984,marscher2006}, 
or mass-loaded `winds' emanating from the accretion disc 
\citep{weymann1991,turnermiller2009}. 
Outflows in AGN offer a 
potential feedback mechanism through which the central source can 
affect its environment \citep{king2003,king2005,fabian2012}
-- feedback that is required in models of galaxy evolution \citep{springel2005}
and may explain the `$M-\sigma$' relation \citep{silkrees1998,haring2004}.

@@

{\bf ksl: I don't see what this paragraph has to do with our problem, which we need to clearly explain} Despite the clear importance of ADWs in understanding AGN SEDs and accretion physics, 
much of the underlying outflow physics remains highly uncertain. 
Several possible driving mechanisms for ADWs have been proposed, including
thermal pressure \citep{weymann1982, begelman1991}, magnetocentrifugal forces 
\citep{blandfordpayne,pelletier_pudritz} and 
radiation pressure on spectral lines \citep[`line-driving';][]{lucysolomon1970,shlosman1985,MCGV95}.
Of these, line-driving is possibly the most attractive, as
strong absorption lines are already seen in BALQSOs and the X-ray spectra of AGN 
\citep{reeves2003,poundsreeves2009,tombesi2010a}.
The presence of line-locked features \citep{bowler2014} 
and the `ghost of \la' (Arav et al. 1996; Arav 1996; North 2006; but see 
also Cottis et al. 2010) \nocite{arav1995, arav1996, north2006,cottis2010}
in the spectra of some BALQSOs also gives clearer evidence that line-driving is
at least partially contributing to the acceleration of the wind.

@@

We have been engaged in a project to determine whether it is possible to simulate the properties of the spectra of AGN, including BAL QSOs, using simple kinematic prescriptions for biconical disc winds using a Monte Carlo radiative transfer code that calculates the ionization structure of the wind and simulates the spectra from such a system \citep[][hereafter H13]{simlong2008,sim2010,higginbottom2013}.   The results have encouraging in the sense that in H13, we showed we could produce simulated spectra that resembled that of BAL QSOs, as long as the luminosity of the central engine was relatively low, of order \POW{43}{\LUM} and the mass loss rate was relatively high, of order the mass accretion rate.  However, at higher luminosities, the wind was so ionized that UV absorption lines were not produced.  In addition, and in part due to limitations in our radiative transfer code, the model failed to produce spectra with strong emission lines at any inclination angle.

Here we attempt to address both of these issues, by introducing clumping into our model and a more complete treatment of H and He into our radiative transfer calculations.   The remainder of this paper is organized as follows:
% There has been some success using simple kinematic prescriptions for biconical disc winds to model  AGN and quasar outflows \citep[][hereafter H13]{simlong2008,sim2010,higginbottom2013}.  H13 successfully produced a benchmark model for BALQSOs. However, the model had two key drawbacks. First, an unrealistically low X-ray luminosity was required in order to prevent over-ionization of the outflow. Second, the model failed to produce the strong emission lines  required at low inclinations in a unified model. In this paper, we attempt to address these issues, and test the disc wind  unification model using Monte Carlo radiative transfer (MCRT) and photoionization  calculations. The paper is organised as follows.
In section 2, we describe some of the important photoionization 
and MCRT aspects of the code. In section 3, we outline the model, including 
a description of our clumping implementation. In section 4, we present the results 
from a clumped model. In section 5 we discuss our results, 
focussing particular on the anisotropy of 
disc emission and GR effects, and finally, in section 6, we summarise our findings.

@@

For this study, we use the MCRT code \py\  we have developed to carry out our radiative transfer and photoionization
calculations in non-local thermodynamic equilibrium (non-LTE) in a variety of disk-wind systems. 
We have used the code  to model accreting white dwarfs 
(Long \& Knigge 2002, hereafter LK02; Noebauer et al. 2010; 
Matthews et al. 2015, hereafter M15), young-stellar objects \citep{simmacro2005}
and quasars (H13).\nocite{noebauer, M15, LK02}  LK02 provides the basic description of the original code; improvements, see especially H14 and M15, have been described in subsequent reports.  A full description of the code will be published elsewhere; here we focus on the changes made to improve the ionization calculation and to allow for clumping in the wind.
%These studies contain extensive detail on the code,  so we only briefly describe the key elements of the global ionization calculation and other important aspects.

{\bf ksl - I a a bit torn over whether we add a few 3 sentences to describe the basic idea.  Some will be reading this for the first time.  Items that one would include are that we calculate the radiative transfer in a cylindrical coordinate system of dimension ?? x ??.  We have ionization and spectral cycles. In each cycle, ?? photon bundles are generated. We stop the ionization cycles after they converge, without any details.  I think this would be useful because we introduce terms like ``cell' later on.', and without some kind of description they have no context}

@@

Our approach to line transfer is based upon the macro-atom implementation developed by \cite{lucy2002, lucy2003}. In our case, for reasons of computational effieciency, we adopt the  hybrid macro-atom scheme 
described by M15 (an improvement over that used by H14 for our earlier quasar study), 
in which the energy flows
through the system are described in terms of indivisible
energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section 2.3).
These energy packets interact with either two-level `simple ions'
or full `macro-atoms'. 
Our scheme allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements which are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions (see M15). In this study,
only H \& He are treated as macro-atoms, because 
we expect recombination to be important
in determining their level populations and resultant line emission, and because we were especially interested in the contribution to AGN spectra of \LA.  In H14, all atoms were treated in a two-level approximation.

@@

Macro-atoms have their ion and level populations derived from
MC rate estimators as described by M15. Previously (LK02, H13, M15),
we used a modified Saha approach to calculate the ionization fractions
of simple-ions. As part of  this effort, we have now improved {\sc PYTHON} to explicitly solve the 
rate equations between ions in non-LTE. 
This dispenses with a number of small assumptions made in the modified Saha approach, 
is more numerically stable, 
and, in principle, allows the direct addition of extra physical 
processes such as Auger ionization.

{\bf ksl - Do we need the rest of this, or could it be postponed until a detailed paper on the program }

%To avoid approximating the complex AGN SED as e.g. a dilute blackbody \citep[][M15]{ML93},
We model the SED in a grid cell using the technique described by H13. In this scheme,
the mean intensity in a series of $n$ bands is modeled as a normalised power law in 
frequency $\nu$
\begin{equation}
J_{\nu,i}=K_{pl}\nu^{\alpha_{pl}},
\end{equation}
for a band $i$, or an exponential 
\begin{equation}
J_{\nu,i}=K_{exp} e^{(-h\nu/k T_{exp})}.
\end{equation}
Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
deduced from the band-limited radiation field estimators.
The ionization rate out of ion $j$ can then be written as 
\begin{equation}
R_{j,j+1}(J)= 
\displaystyle{
n_i \left(C_{j} n_e + 
\sum_{band~i=0}^{n}~
{\int_{\nu_i}^{~\nu_i+1}{J_{\nu,i}\sigma_j(\nu)\nu^{-1}d\nu}}
\right),}
\end{equation}
where $\sigma_j$ is the photoionization cross-section and $C_{j}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $j$ is given by 
\begin{equation}
R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{DR} + \alpha^j_{CR}) n_{j+1} n_e,
\end{equation}
where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
The subscripts denote radiative, dielectronic and 
collisional (three-body) recombination, respectively.
We neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann equation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$.

@@

{\bf This needs an introductory paragraph.  Should state our earlier modeling of AGN H13 was based on a smooth outflow.  It should reference the introduction where we say there are observational indications of clumping.  Should say that the difficulty associated with clumping is two fold - computationally implies many grid cells,  many potential parameters.  Ultimately, say that for simplicity we use microclumping.  Then say this has been used elsewhere, but with more recent references than what we have.  The point is we need to motivate micro clumping before we say what it means to the code  It's all mixed up as written. Aside: at some point this paper should mention H14, and hydro models more generally}

@@

{\bf We need to describe what we did before getting to results.  We started with parameters similar to those of H14, we created a small grid.  Limits of the grid are given in table 1.  This should be a paragraph or section before this one.  Not all quasars have to have the same parameters and so we have to explain that we have elected to describe one model in detail.  Was our goal to fix the luminosity and see what was required.

In next sections, we really need to deal with the difference between this class of models and "the model", and especially "our model".  We may need to at least in words discuss whether we have "fine tuned" or not, by saying something about other models.}

@@

\subsection{Synthetic Spectra: Comparison to Observations}

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/windnew3.png}
\caption
{
A cartoon showing the broad classes of sightlines 
in our model, illustrates how geometric effects lead to 
the different emergent spectra. The colour gradient is approximate,
but indicates the stratified ionization structure, 
from highly ionized (yellow) to low ionization (purple) material.
%[FIGURE NEEDS IMPROVING]
}
\label{fig:sightline}
\end{figure}

@@

Figure~\ref{fig:uvspec} shows the synthetic spectrum in the UV from our model. 
We also show a comparison to composite quasar and BALQSO spectra.
We show a cartoon illustrating how geometric effects determine
the output spectra in figure~\ref{fig:sightline}.  {\bf - Need to explain the source of the BAL composites}

@@

We find that the model can produce significant line emission
at low inclinations, particular in \civ, 
and the improved treatment of recombination results in a strong \la\
line. In the context of unification, this is a promising result, 
and shows that a biconical wind can produce significant 
emission at `quasar-like' angles. To assess the ability of the model to match real 
quasar spectra, we also show {\sl Sloan Digital Sky Survey} (SDSS) quasar composites from
\cite{reichard2003}, normalised to the flux at 2000\AA\ in each panel. Our simulated spectra do not contain
the semi-forbidden 
intercombination lines seen in quasar spectra because our radiative transfer code 
currently does not have a treatment for semi-forbidden lines.  {\bf Isn't this just a statement about the fact that we have not implemented CIII as a full macro-atom; if so we should say that}
This is especially noticeable with the strong C~\textsc{iii}]~1909\AA\, 
line in the quasar composite spectra.  
The critical density of the C~\textsc{iii}]~1909\AA\, line 
is $n_e\sim10^{9.5}~\rm{cm^{-3}}$, which is higher than much of the 
outer portion of our wind. We therefore expect intercombination lines
to become important coolants in the outer portion of the wind. {\bf we have gone from their being lines we should see, to important coolants.  I think we should stick to the fact that we would think our code would see them if we put C in as a macro atom.  Also I think this who discussion of CIII should be after we have discussed what looks right with the emission lines.}

@@

{\bf: I am not convinced this in the right place in the paper, because here we explore an alternative hypythesis.  It might go better at the end.  But a good many of the following sections are not that well motivated.}

@@

When comparing BALQSO and quasar composites, it is apparent
that they possess remarkably similar line strengths and widths 
\citep[e.g.][]{reichard2003}.  {\bf I think we mean EW here; otherwise I don't understand what we are talking about.  We could say ``line shapes and equivalent widths''.  I also think this requries more than one sentence of introduction.}
This presents a challenge to our model, as well as the geometric 
unification picture in general, because these quantities should vary with inclination angle.
Limb darkening and foreshortening causes the disc continuum emission to be 
strongly anisotropic, yet the line emission in our models is much more isotropic. 
This has the effect of enhancing the line-to-continuum ratios at high inclinations. 
To construct a scenario where emission 
line equivalent widths are comparable at all inclinations requires 
significant fine-tuning when considering a classical thin disc. {\bf Is the previous sentence a throwaway line, or has some person done this.  If so, please reference; if not please delete.}
%An alternative, simpler solution is that the emergent continuum is roughly isotropic.

Obviously,  one possible solution to this problem is hypothesize a more isotropic distribution for the emergent condition than predicted by a ``classical'' disc. 
General relativistic effects -- specifically, light bending
and relativistic beaming -- can cause 
the accretion disc SED to become more isotropic \citep[e.g.][]{zhang1997,munozdarias2013}.  Incorporating these effects into {\sc PYTHON} would be a very difficult and time-consuming challenge. Therefore, for the purposes of this discussion we have chosen to investigate whether GR effects could produce enough isotropy to address the EW problem, we have generated 
\ GR disc spectra, we use the code \agn\ \citep{hubeny2000,davishubeny2006,davis2007}. 
The output flux at three different wavelengths
as a function of inclination for an \agn\ model with the same disc and BH parameters
as our clumpy wind model is shown in figure~\ref{fig:f2000}.  The results indicate that 
the effects of GR on an AGN disc are much less extreme 
in the UV portion of the spectrum than the calculations
by \cite{zhang1997} for the X-rays in X-ray binaries.
The emission for the disks of AGN is still strongly anisotropic.
GR alone therefore cannot explain the line ratio trends in quasars by making the disc
emit more isotropically.
This supports the findings of \cite{risaliti2011}, who find that EW distributions 
in quasars are consistent with anisotropic emission from optically thick, 
disc-like structures for {\em both} the continuum source and BLR. 
We will explore these ideas further in a future study.  {\bf Wait!  All this says is that there has to be an alternative explanation.  It is not support for any specific explanation!. This last few sentences should be a separate paragraph.  They should start out: An alernative explanation proposed by risalite is that...}

@@

{\bf:  Again this section needs an introduction.  Why is this important?}

@@

{\bf Note: One of the things I definitely wondered about is whether our "model" would produce the fraction of BALs seen in the sky}

### 2015-09-01 paper_mnras.tex
Knox's first set of changes, trying to commit them again

%\author[Matthews et al.]
\author[Matthews et al.]{
\parbox[t]{\textwidth}{
J.~H.~Matthews$^1$\thanks{jm8g08@soton.ac.uk}, C.~Knigge$^1$,
N.~Higginbottom$^1$, K.~S.~Long$^2$, S.~A.~Sim$^3$ and S.~W.~Mangham$^1$
}
\medskip  
\\$^1$School of Physics and Astronomy, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom
\\$^2$Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD, 21218
\\$^3$School of Mathematics and Physics, Queens University Belfast, University Road, Belfast, BT7 1NN, Northern Ireland, UK
}

>>>>

%\author[Matthews et al.]
\author[Matthews et al.]{
\parbox[t]{\textwidth}{
James~H.~Matthews$^1$\thanks{jm8g08@soton.ac.uk}, Christian~Knigge$^1$,
Nick~Higginbottom$^1$, Knox~S.~Long$^2$, Stuart~A.~Sim$^3$ and Sam~W.~Mangham$^1$
}
\medskip  
\\$^1$School of Physics and Astronomy, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom
\\$^2$Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD, 21218
\\$^3$School of Mathematics and Physics, Queens University Belfast, University Road, Belfast, BT7 1NN, Northern Ireland, UK
}
